# Exploratory Analysis {#exploratory}

## What data is

Data come packed from the API end point /complete in JSON format. A preliminary data assessment evidences 34 covariates and a  250 rows, even though the number could be varyign accoriding to the API query sent. Json file is then parsed by the convenient read_csv from the readr package mutating columns by their respective type. 
below the explanation of the columns:

ID: _numeric_ a primary key number from immobiliare to identify univocally the advertisement 
LAT: _numeric_ Exact latitude where the apartement is located (that compose the spatial covariate) 
LONG: _numeric_ Exact longitude where the apartement is located (that compose the spatial covariate)
CONDOM:  _numeric_ The condominium price which is an important component in the final cost structure
BUILDAGE: _numeric_ the year in which the building has been built
FLOOR: _numeric_ the floor at which is located the apartement/ house
INDISVAP: _categorical_ If it the rental ad regards an indipendent structure or an apartement in a building
...




At a first glance it could be possible to see some missing values in some consistent measure in LOWPRCE and its derivations, Energy calss Disp Pauto. 

```{r, echo=FALSE, message=FALSE }
library(readr)
library(tidyverse)
  data =  read.csv("data/data.csv")
```

## Data Glimpse 
```{r}
library(knitr)
glimpse(data) 
```



## Explorative Analysis

###  Semivariogram Covariogram

