# Exploratory Analysis {#exploratory}

## What data is

Data come packed from the API end point /complete in JSON format. A preliminary data assessment evidences 34 covariates and a  250 rows, even though the number could be varyign accoriding to the API query sent. Json file is then parsed by the convenient read_csv from the readr package mutating columns by their respective type. 
below the explanation of the columns:

ID: _numeric_ a primary key number from immobiliare to identify univocally the advertisement 
LAT: _numeric_ Exact latitude where the apartement is located (that compose the spatial covariate) 
LONG: _numeric_ Exact longitude where the apartement is located (that compose the spatial covariate)
CONDOM:  _numeric_ The condominium price which is an important component in the final cost structure
BUILDAGE: _numeric_ the year in which the building has been built
FLOOR: _numeric_ the floor at which is located the apartement/ house
INDISVAP: _categorical_ If it the rental ad regards an indipendent structure or an apartement in a building
...




At a first glance it could be possible to see some missing values in some consistent measure in LOWPRCE and its derivations, Energy calss Disp Pauto. 

```{r, echo=FALSE, message=FALSE }
library(readr)
library(tidyverse)
  data =  read.csv("data/data.csv")
```

## Data Glimpse 
```{r}
library(knitr)
glimpse(data) 
```



We use the dataset extracted from the “PERVAL” database. This database
contains all information on property transactions, property attributes as well
as buyers and sellers profiles. Additionally, most observations in the database
are geo-referenced and therefore spatial analysis is used.
After removing the apartments with missing characteristics and eliminating
the apartments corresponding to the tails11 of the price distribution, there are
7634 observations spanning from 2006 to 2017.
The key variables of the dataset include the sale price, apartment structural
characteristics and accessibility variables. The description of all variables is
listed in Table 1.
The summary statistics of all continuous variables are presented in Table



## Explorative Analysis

###  Semivariogram Covariogram

