# Model Selection & Fitting{#modelspec}

```{r, include=FALSE}
knitr::opts_chunk$set(
  warning = FALSE,  
  strip.white = TRUE,
  message = FALSE,
  cache = FALSE,
  echo = FALSE
)

# libraries are in libs.R
# source(here::here("libs.R"))

library(geoR, warn.conflicts = F, quietly = T )
library(brinla, warn.conflicts = F, quietly = T )
library(ggplot2, warn.conflicts = F, quietly = T )
library(tibble, warn.conflicts = F, quietly = T )
library(knitr, warn.conflicts = F, quietly = T)
library(latex2exp, warn.conflicts = F, quietly = T)
library(faraway, quietly = T, warn.conflicts = F)
library(patchwork, warn.conflicts = F, quietly = T)
library(INLA, warn.c, warn.conflicts = F, quietly = T)
library(rgdal, warn.conflicts = F, quietly = T)
library(ggthemes, warn.conflicts = F, quietly = T)
library(sf, warn.conflicts = F, quietly = T)

theme_nicco = function (base_size = 11, base_family = "") {
  theme_bw() %+replace% 
    theme(
      text = element_text(family = "sans", size = 12),
      plot.title = element_text(face = "bold", size = 14, margin=margin(0,0,30,0)),
      panel.background  = element_blank(),
      axis.ticks = element_line(colour = "grey70", size = 0.2),
      plot.background = element_rect(fill="white", colour=NA),
      panel.border = element_rect(linetype = "blank", fill = NA),
      legend.background = element_rect(fill="transparent", colour=NA),
      legend.key = element_rect(fill="transparent", colour=NA)
    )
}
```




- metti @Ling che ha messo i prioris come defaults nella inla call


- link tanta roba con inlabru [link](https://www.muscardinus.be/2018/07/inlabru-bru/)


[ref.](https://inbo.github.io/tutorials/tutorials/r_inla/spatial.pdf)
```{r, eval=FALSE}
library(readr)
data = read_csv("data/data2021.csv")
mesh2 <- inla.mesh.2d(confini, max.edge = c(0.2, 0.2),cutoff = 3)
ggplot(data) + gg(mesh2) + geom_sf() +
ggtitle(paste("Vertices: ", mesh2$n)) + coord_sf(datum = st_crs(5880))

```


fit with the suggestion of INLA in [ref](https://www.r-inla.org/faq#h.sxjo232d6ho5)
```{r, eval=F}
library(spdep)
library(spDataLarge)
border <- readShapePoly(system.file("data/confini/A090101_ComuneMilano.shp", package="spdep")[1])



```




```{r, eval=F}

confini = readOGR(dsn = "data/confini/A090101_ComuneMilano.shp") 
  # st_as_sf()
proj4string(confini)
confini_proj = spTransform(confini, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))%>% 
                              st_as_sf()
plot(confini_proj)                     
ggplot(NULL) +
  geom_sf(data = confini_proj)+
  geom_point(data = dati, aes(x = long, y = lat), size = 4, shape = 23, fill = "darkred") +
  theme_map() 
```


```{r, eval=F}
require(sp)  # package to work with spatial data
require(rgdal)
library(leaflet)
library(sf)


milan_borders = readOGR("data/confini/A090101_ComuneMilano.dbf")
proj4string(milan_borders)
milan_borders_pj <- spTransform(milan_borders, CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
pal = colorBin("viridis", bins = c(100,200,300,400))

leaflet(dati) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircles(lng = ~long, lat = ~lat, color = ~pal(price))%>%
  addLegend("bottomright",
    pal = pal, values = ~price,
    title = "Price"
  ) %>%
  addScaleBar(position = c("bottomleft"))

```





[mega ref](https://inlabru-org.github.io/inlabru/)

[mega mega ref](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.13168)

```{r, eval=FALSE}
library(inlabru)
formula = price ~ condom + totlocali + sqfeet 
```   



```{r, eval=FALSE}
ggplot(data)+
  geom_point(aes(lat, long), size = 2) +
  coord_fixed(ratio = 1) +
  scale_color_gradient(low = "blue", high = "orange") +
  geom_sf(data = confini) +
  theme_bw()
```






**covariate choice **
## Model Selection

- wang pagina 49
In regression analysis, we often want to find a reduced model with the best subset of
the variables from the full model. The model selection in frequentist analysis is com-
monly based on Akaike information criterion (AIC), a MLE-based criterion. Back to
the air pollution data example, a stepwise model selection procedure using AIC can
be implemented by the function stepAIC in R library MASS

```{r, eval=FALSE}
library(MASS)
usair.step <- stepAIC(usair.lm1, trace = FALSE)
usair.step$anova
```


## Mesh Construction 200 blangiardo



how to fit a spatial model
https://www.r-inla.org/faq#h.sxjo232d6ho5


